### 2020-08-28 Manuscript/What.lies.ahead.reviewed.tex
Started updating the ms

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}
\usepackage{xcolor}

@@

Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database (CITE CRUX). We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. The multivariate tests (PERMANOVA, CAP, k-means clustering) used these index values as the species abundances and Bray-Curtis as the dissimilarity measurement between samples.

@@

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.   {\color{red}[ADD DETAIL HERE]}

%We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer''). To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017. See Supporting Information for details of sampling parameters.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. {\color{red} Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models.} We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}. Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (Red) and San Juan Island (Blue). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as facets. {\color{red} CHANGE y-scale for Dinophyceae}}}
\label{fig:Fig1}
\end{figure}
%%%%%

@@

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F-model = 15.4; p $<$ 0.001). {\color{red} RPK: still up-to-date? no adonis code in the Rmd file that I saw} . We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2 shows results for Hood Canal in summer as an example; full analysis in the Supporting Information). Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3 or 4) to identify groups of samples that were similar to one another with respect to biological community. Each biological cluster (colored hulls, Fig. 2) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

{\color{red} RPK: I think everything below needs updating, according to some final analysis}

These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal in summer for example (Fig. 2C), the colder water communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp. and \textit{Chaetoceros} sp. Warmer waters were dominated by \textit{Bathycoccus prasinos} and \textit{Cymbomonas} sp., while at lower pH values, marine communities showed \textit{Litchtheimia ramosa}, \textit{Micromonas pusilla} at higher relative abundances.

@@

The low-pH planktonic communities (cluster 3, Fig. 2) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where cluster 3 is the most likely community -- to occur more than two-thirds of the time (68\%).

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2020-08-12.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

@@

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-12.Other.sel.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

>>>>

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}
\usepackage{xcolor}
\usepackage{soul}

@@

Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database \cite{curd2019anacapa}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.

\subsection*{Present scenario community analyses}

{\color{red}
We performed a series of multivariate tests to investigate the community changes across the environmental space we observed. We used the index of abundance as the species abundances and Bray-Curtis as the dissimilarity measurement between samples. A global PERMANOVA test (Bray-Curtis ~ Area) further justified the subsequent separated  analysis of the Hood Canal and San Juan Island samples; a Constrained Analysis of Principal Components (CAP, \cite{Anderson}) investigated the distinctiveness of plankton communities across the gradient of temperature, pH and salinity.

Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community. The SIMPER procedure from the r package \textit{vegan} revealed the taxa most contributing to between-cluster differences.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models. We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.
}

@@

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.   {\color{red}We fit our current environmental observations to a normal distribution, and used the parameters of that distribution alongside a linear variation of temperature and pH between the present and the 2095 scenario to generate 1000 simulations of each year scenario. The scenario labelled as 2095 is the result of 5 simulations for the years 2091-2095. We removed outliers from these simulations by keeping the 95\% percentile of observations. See the R code in the supplementary information to follow the process of generating these datasets (lines 98-135).
%We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer''). To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017. See Supporting Information for details of sampling parameters.

\st{For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables.   We calculated the probability of each community, given these predictors, using the package `nnet` . Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year. }}

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal ({\color{red}dark gold}) and San Juan Island ({\color{red}turquoise}). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.{\color{red}\textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately} \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as facets. {\color{red} CHANGE y-scale for Dinophyceae DONE}}}
\label{fig:Fig1}
\end{figure}
%%%%%

@@

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F-model = 5.4; p $<$ 0.0001). {\color{red} RPK: still up-to-date? no adonis code in the Rmd file that I saw. RGS: that's true - I added new chunk of code line 382 \& updated results}. We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2 shows results for Hood Canal; full analysis in the Supporting Information). 
{\color{red}\st{Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.}}
 Each biological cluster (colored hulls, Fig. 2.A) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

{\color{red} RPK: I think everything below needs updating, according to some final analysis. TODO: RGS}

These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal for example (Fig. 2C), the colder water and higher pH communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp. and \textit{Chaetoceros} sp. Warmer waters were dominated by \textit{Bathycoccus prasinos} and \textit{Cymbomonas} sp., while at lower pH values, marine communities showed \textit{Litchtheimia ramosa}, \textit{Micromonas pusilla} at higher relative abundances.

@@

The low-pH and {\color{red}high-temperature} planktonic communities (cluster 3, Fig. 2{\color{red}A and B}) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where cluster 3 is the most likely community -- to occur more than two-thirds of the time (68\%).

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

@@

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-27.Other.sel.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}
